---
title: "Thesis Industry C14"
author: "Michael Battaile"
date: "8/2/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Load library and import data
```{r echo = T, message = F, results = 'hide'}
library(ranger)
library(rpart)
library(rpart.plot)
library(dplyr)
library(caret)
library(Rcpp)
library(xgboost)
library(class)
library(VIM)
library(GGally)
library(corrplot)
library(stargazer)
library(data.table)
library(moments)
```


**Load Data**
```{r}
#load and import data

data <- read.csv(file = "data_inflation_adjusted.csv", fileEncoding="UTF-8-BOM")

macro_variables <- read.csv(file = "macro_indicators_romania.csv", fileEncoding="UTF-8-BOM")

data <- inner_join(data, macro_variables, by="year")

data <- data[data$nace_division == 14,]

#set seed

set.seed(123)
```


**Convert Data Types**
```{r}
#convert data to correct data types

data$reg_number <- as.integer(data$reg_number)
data$Default <- as.factor(data$Default)
data$nace_code <- as.factor(data$nace_code)
data$nace_division <- as.factor(data$nace_division)
data$nace_section <- as.factor(data$nace_section)
data$year <- as.numeric(data$year)
data$noncurrent_assets <- as.numeric(data$noncurrent_assets)
data$current_assets <- as.numeric(data$current_assets)
data$inventories <- as.numeric(data$inventories)
data$receivables <- as.numeric(data$receivables)
data$cash_and_bank <- as.numeric(data$receivables)
data$prepaid_expenses <- as.numeric(data$prepaid_expenses)
data$payables <- as.numeric(data$payables)
data$advance_payments <- as.numeric(data$advance_payments)
data$provisions <- as.numeric(data$provisions)
data$equity <- as.numeric(data$equity)
data$share_capital <- as.numeric(data$share_capital)
data$state_owned_capital <- as.numeric(data$state_owned_capital)
data$public_owned_capital <- as.numeric(data$public_owned_capital)
data$net_turnover <- as.numeric(data$net_turnover)
data$total_income <- as.numeric(data$total_income)
data$total_expenses <- as.numeric(data$total_expenses)
data$net_profit_or_loss <- as.numeric(data$net_profit_or_loss)
data$gross_profit_or_loss <- as.numeric(data$gross_profit_or_loss)
data$employee_count <- as.numeric(data$employee_count)
data$gdp_growth <- as.numeric(data$gdp_growth)
data$inflation <- as.numeric(data$inflation)
data$unemployment <- as.numeric(data$unemployment)
data$interest <- as.numeric(data$interest)
```


**Convert 0 and NAN to NA**
```{r}
#turn 0s and NAN into NA

data[data == 0] <- NA

is.nan.data.frame <- function(x)
  do.call(cbind, lapply(x, is.nan))

data[is.nan(data)] <- NA
```


**Add YoY Columns**
```{r}
#add yoy columns

longitudinal <- data %>%
  group_by(reg_number) %>% 
  arrange(year, .by_group = TRUE) %>%
  mutate(current_assets_yoy = (current_assets/lag(current_assets) - 1) * 100)%>%
  mutate(receivables_yoy = (receivables/lag(receivables) - 1) * 100) %>%
  mutate(cash_and_bank_yoy = (cash_and_bank/lag(cash_and_bank) - 1) * 100) %>%
  mutate(payables_yoy = (payables/lag(payables) - 1) * 100) %>%
  mutate(equity_yoy = (equity/lag(equity) - 1) * 100) %>%
  mutate(share_capital_yoy = (share_capital/lag(share_capital) - 1) * 100) %>%
  mutate(total_expenses_yoy = (total_expenses/lag(total_expenses) - 1) * 100) %>%
  mutate(gross_profit_or_loss_yoy = (gross_profit_or_loss/lag(gross_profit_or_loss) - 1) * 100) %>%
  mutate(net_profit_or_loss_yoy = (net_profit_or_loss/lag(net_profit_or_loss) - 1) * 100)
```


**Delete First Year Observations**
```{r}
#delete first years

universal_unclean <- longitudinal %>%
  group_by(reg_number) %>%
  filter(row_number() > 1)

summary(universal_unclean)
```


**Reduce Variables and Observations w/o Enough Data**
```{r}
#reduce variables by 20% or less NAs

universal <- universal_unclean[,which(colMeans(!is.na(universal_unclean)) > 0.8)]

#reduce datapoints by 20% or less NAs

universal <- universal[which(rowMeans(!is.na(universal)) > 0.8),]

summary(universal)
```


**Compute and Check Mean/MEdian Ratio to Judge Skew**
```{r}
#compute and check mean/median ratios to judge skew and ideal average for imputation

medians <- c(median(universal$current_assets, na.rm = TRUE),
             median(universal$receivables, na.rm = TRUE),
             median(universal$cash_and_bank, na.rm = TRUE),
             median(universal$payables, na.rm = TRUE),
             median(universal$equity, na.rm = TRUE),
             median(universal$share_capital, na.rm = TRUE),
             median(universal$total_expenses, na.rm = TRUE),
             median(universal$gross_profit_or_loss, na.rm = TRUE),
             median(universal$net_profit_or_loss, na.rm = TRUE),
             median(universal$current_assets_yoy, na.rm = TRUE),
             median(universal$payables_yoy, na.rm = TRUE),
             median(universal$equity_yoy, na.rm = TRUE),
             median(universal$share_capital_yoy, na.rm = TRUE),
             median(universal$total_expenses_yoy, na.rm = TRUE),
             median(universal$gross_profit_or_loss_yoy, na.rm = TRUE),
             median(universal$net_profit_or_loss_yoy, na.rm = TRUE))

trimmed_means <- c(mean(universal$current_assets, na.rm = TRUE),
                   mean(universal$receivables, na.rm = TRUE),
                   mean(universal$cash_and_bank, na.rm = TRUE),
                   mean(universal$payables, na.rm = TRUE),
                   mean(universal$equity, na.rm = TRUE),
                   mean(universal$share_capital, na.rm = TRUE),
                   mean(universal$total_expenses, na.rm = TRUE),
                   mean(universal$gross_profit_or_loss, na.rm = TRUE),
                   mean(universal$net_profit_or_loss, na.rm = TRUE),
                   mean(universal$current_assets_yoy, na.rm = TRUE),
                   mean(universal$payables_yoy, na.rm = TRUE),
                   mean(universal$equity_yoy, na.rm = TRUE),
                   mean(universal$share_capital_yoy, na.rm = TRUE),
                   mean(universal$total_expenses_yoy, na.rm = TRUE),
                   mean(universal$gross_profit_or_loss_yoy, na.rm = TRUE),
                   mean(universal$net_profit_or_loss_yoy, na.rm = TRUE))

standard <- c(sd(universal$current_assets, na.rm = TRUE),
             sd(universal$receivables, na.rm = TRUE),
             sd(universal$cash_and_bank, na.rm = TRUE),
             sd(universal$payables, na.rm = TRUE),
             sd(universal$equity, na.rm = TRUE),
             sd(universal$share_capital, na.rm = TRUE),
             sd(universal$total_expenses, na.rm = TRUE),
             sd(universal$gross_profit_or_loss, na.rm = TRUE),
             sd(universal$net_profit_or_loss, na.rm = TRUE),
             sd(universal$current_assets_yoy, na.rm = TRUE),
             sd(universal$payables_yoy, na.rm = TRUE),
             sd(universal$equity_yoy, na.rm = TRUE),
             sd(universal$share_capital_yoy, na.rm = TRUE),
             sd(universal$total_expenses_yoy, na.rm = TRUE),
             sd(universal$gross_profit_or_loss_yoy, na.rm = TRUE),
             sd(universal$net_profit_or_loss_yoy, na.rm = TRUE))

skewness <- 3 * (trimmed_means - medians) / standard

names(skewness) <- c("current_assets", "receivables", "cash_and_bank", "payables", "equity", "share_capital", "total_expenses", "gross_profit_or_loss", "net_profit_or_loss", "current_assets_yoy", "payables_yoy", "equity_yoy", "share_capital_yoy", "total_expenses_yoy", "gross_profit_or_loss_yoy", "net_profit_or_loss_yoy")

print(skewness)
```


**NA Imputation**
```{r}
#NA imputation

universal$current_assets[is.na(universal$current_assets)] <- mean(universal$current_assets, na.rm = TRUE)
universal$receivables[is.na(universal$receivables)] <- mean(universal$receivables, na.rm = TRUE)
universal$cash_and_bank[is.na(universal$cash_and_bank)] <- mean(universal$cash_and_bank, na.rm = TRUE)
universal$payables[is.na(universal$payables)] <- mean(universal$payables, na.rm = TRUE)
universal$equity[is.na(universal$equity)] <- mean(universal$equity, na.rm = TRUE)
universal$share_capital[is.na(universal$share_capital)] <- mean(universal$share_capital, na.rm = TRUE)
universal$total_expenses[is.na(universal$total_expenses)] <- mean(universal$total_expenses, na.rm = TRUE)
universal$gross_profit_or_loss[is.na(universal$gross_profit_or_loss)] <- mean(universal$gross_profit_or_loss, na.rm = TRUE)
universal$net_profit_or_loss[is.na(universal$net_profit_or_loss)] <- mean(universal$net_profit_or_loss, na.rm = TRUE)
universal$current_assets_yoy[is.na(universal$current_assets_yoy)] <- mean(universal$current_assets_yoy, na.rm = TRUE)
universal$receivables_yoy[is.na(universal$receivables_yoy)] <- median(universal$receivables_yoy, na.rm = TRUE)
universal$cash_and_bank_yoy[is.na(universal$cash_and_bank_yoy)] <- median(universal$cash_and_bank_yoy, na.rm = TRUE)
universal$payables_yoy[is.na(universal$payables_yoy)] <- mean(universal$payables_yoy, na.rm = TRUE)
universal$equity_yoy[is.na(universal$equity_yoy)] <- mean(universal$equity_yoy, na.rm = TRUE)
universal$share_capital_yoy[is.na(universal$share_capital_yoy)] <- mean(universal$share_capital_yoy, na.rm = TRUE)
universal$total_expenses_yoy[is.na(universal$total_expenses_yoy)] <- mean(universal$total_expenses_yoy, na.rm = TRUE)
universal$gross_profit_or_loss_yoy[is.na(universal$gross_profit_or_loss_yoy)] <- mean(universal$gross_profit_or_loss_yoy, trim = .025, na.rm = TRUE)
universal$net_profit_or_loss_yoy[is.na(universal$net_profit_or_loss_yoy)] <- mean(universal$net_profit_or_loss_yoy, trim = .025, na.rm = TRUE)


universal1 <- universal[c(1,2,3,4,5,15,6,7,8,9,10,11,12,13,14,20,21,22,23,24,25,26,16,17,18,19)]
p <- median(universal$receivables_yoy, na.rm = TRUE)
summary(universal1)
```


**Check Correlations for Colinearity**
```{r}
#check for strong correlations

universal_numeric <- universal1[7:26]

scaled_data <- as.data.frame(scale(universal_numeric))

cor <- cor(scaled_data, use = "complete.obs")

corrplot(cor, 
         method = "number", 
         type = "upper", 
         order = "hclust", # reorder by the size of the correlation coefficients
         tl.cex = 0.5, # font size of the variable labels
         tl.col = "black", # color of the variable labels
         tl.srt = 45, # rotation angle for the variable labels
         number.cex = 0.5 # font size of the coefficients
         )
```


**Remove Highly Correlated Variables**
```{r}
#remove correlated variables

reg_data_unclean <- c(universal1[c(6,7,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,26)])


reg_data_unclean <- data.frame(reg_data_unclean)

summary(reg_data_unclean)
```


**Check for Outliers**
```{r}
#check for outliers via boxplot

boxplot <- boxplot(reg_data_unclean)
```


**Remove Outliers**
```{r}
#remove outliers

Q <- quantile(reg_data_unclean$current_assets, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean$current_assets)
reg_data_unclean1 <- subset(reg_data_unclean, reg_data_unclean$current_assets > (Q[1] - 1.5*iqr) & reg_data_unclean$current_assets < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean1$payables, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean1$payables)
reg_data_unclean <- subset(reg_data_unclean1, reg_data_unclean1$payables > (Q[1] - 1.5*iqr) & reg_data_unclean1$payables < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean$share_capital, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean$share_capital)
reg_data_unclean1 <- subset(reg_data_unclean, reg_data_unclean$share_capital > (Q[1] - 1.5*iqr) & reg_data_unclean$share_capital < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean1$total_expenses_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean1$total_expenses_yoy)
reg_data_unclean <- subset(reg_data_unclean1, reg_data_unclean1$total_expenses_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean1$total_expenses_yoy < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean$equity, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean$equity)
reg_data_unclean1 <- subset(reg_data_unclean, reg_data_unclean$equity > (Q[1] - 1.5*iqr) & reg_data_unclean$equity < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean1$net_profit_or_loss, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean1$net_profit_or_loss)
reg_data_unclean <- subset(reg_data_unclean1, reg_data_unclean1$net_profit_or_loss > (Q[1] - 1.5*iqr) & reg_data_unclean1$net_profit_or_loss < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean$gross_profit_or_loss, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean$gross_profit_or_loss)
reg_data_unclean1 <- subset(reg_data_unclean, reg_data_unclean$gross_profit_or_loss > (Q[1] - 1.5*iqr) & reg_data_unclean$gross_profit_or_loss < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean1$current_assets_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean1$current_assets_yoy)
reg_data_unclean <- subset(reg_data_unclean1, reg_data_unclean1$current_assets_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean1$current_assets_yoy < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean$payables_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean$payables_yoy)
reg_data_unclean1 <- subset(reg_data_unclean, reg_data_unclean$payables_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean$payables_yoy < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean1$equity_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean1$equity_yoy)
reg_data_unclean <- subset(reg_data_unclean1, reg_data_unclean1$equity_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean1$equity_yoy < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean$share_capital_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean$share_capital_yoy)
reg_data_unclean1 <- subset(reg_data_unclean, reg_data_unclean$share_capital_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean$share_capital_yoy < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean1$gross_profit_or_loss_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean1$gross_profit_or_loss_yoy)
reg_data <- subset(reg_data_unclean1, reg_data_unclean1$gross_profit_or_loss_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean1$gross_profit_or_loss_yoy < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean$total_expenses_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean$total_expenses_yoy)
reg_data_unclean1 <- subset(reg_data_unclean, reg_data_unclean$total_expenses_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean$total_expenses_yoy < (Q[2]+1.5*iqr))

Q <- quantile(reg_data_unclean1$net_profit_or_loss_yoy, probs=c(.25, .75), na.rm = FALSE)
iqr <- IQR(reg_data_unclean1$net_profit_or_loss_yoy)
reg_data <- subset(reg_data_unclean1, reg_data_unclean1$net_profit_or_loss_yoy > (Q[1] - 1.5*iqr) & reg_data_unclean1$net_profit_or_loss_yoy < (Q[2]+1.5*iqr))

summary(reg_data)
```


**Take a Sample with Balanced Defaults and Non-Defaults**
```{r}
#take a sample with equal defaults and non-defaults
no_macro <- reg_data[c(1:15)]

defaults <- no_macro[no_macro$Default== "Yes",]
non_defaults <- no_macro[no_macro$Default== "No",]

default_sample <- defaults[sample(nrow(defaults), 57), ]

non_default_sample <- non_defaults[sample(nrow(non_defaults), 114), ]

sample <- rbind(default_sample, non_default_sample)
```



```{r}
#create data partitions

universal_split <- createDataPartition(sample$Default, 
                                p = 0.8, list = F)
universal_train <- sample[universal_split, ]
universal_test <- sample[-universal_split, ]

summary(universal_train)
summary(universal_test)
```


*Logistic Regression**
```{r}
logistic <- glm(Default ~ ., data = sample, family = binomial(logit))

summary(logistic)

logistic_pred <- predict(logistic, universal_test)

train_control <- trainControl(method = "cv", number = 10)

logistic_fit <- train(Default ~ .,
                 data = universal_train,
                 trControl = train_control,
                 method = "glm",
                 family = "binomial")

logistic_fit_pred <- predict(logistic_fit, universal_test)

confusionMatrix(logistic_fit_pred, universal_test$Default)

postResample(pred = logistic_fit_pred, obs = universal_test$Default)

logistic_test <- mutate(universal_test,
                        obs = Default,
                        pred = predict(logistic_fit, newdata = universal_test),
                        predict(logistic_fit, newdata = universal_test, type = "prob"))

twoClassSummary(logistic_test, lev = levels(logistic_test$Default))
```


**Random Forests**
```{r}
tree = rpart(Default ~ ., data = universal_train)

min_cp <- tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]

tree_prune <- prune(tree, cp = min_cp)

tree_test_pred <- predict(tree, universal_test)

randomforest_grid_ranger <- expand.grid(mtry = seq(2, 4, length = 1), 
                                 splitrule = "extratrees",
                                 min.node.size = 2)

oob <- trainControl(method = "oob")

randomforest_tune <- train(Default ~ ., 
                           data = universal_train,
                           method = "ranger",
                           trControl = oob,
                           tuneGrid = randomforest_grid_ranger)

pred_randomforest_tune <- predict(randomforest_tune, universal_test)

confusionMatrix(pred_randomforest_tune, universal_test$Default)

postResample(pred_randomforest_tune, universal_test$Default)
```


**Boosted**
```{r}
library(xgboost)
boosting <- train(Default ~ ., 
                  data = universal_train,
                  method = "xgbTree",
                  trControl = trainControl(method = "cv", number = 6)
)

pred_boosting <- predict(boosting, universal_test)

confusionMatrix(pred_boosting, universal_test$Default)

postResample(pred = pred_boosting, obs = universal_test$Default)

boosting_test <- mutate(universal_test,
                        obs = Default,
                        pred = predict(boosting, newdata = universal_test),
                        predict(boosting, newdata = universal_test, type = "prob"))

twoClassSummary(boosting_test, lev = levels(boosting_test$Default))
```



```{r}
#take a sample with equal defaults and non-defaults

macro <- reg_data

macro_defaults <- macro[macro$Default== "Yes",]
macro_non_defaults <- macro[macro$Default== "No",]

macro_default_sample <- macro_defaults[sample(nrow(macro_defaults), 57), ]

macro_non_default_sample <- macro_non_defaults[sample(nrow(macro_non_defaults), 114), ]

macro_sample <- rbind(macro_default_sample, macro_non_default_sample)
```



```{r}
#create data partitions

macro_split <- createDataPartition(macro_sample$Default, 
                                p = 0.8, list = F)
macro_train <- macro_sample[macro_split, ]
macro_test <- macro_sample[-macro_split, ]

summary(macro_train)
summary(macro_test)
```



```{r}
logistic <- glm(Default ~ ., data = macro_sample, family = binomial(logit))

summary(logistic)

logistic_pred <- predict(logistic, macro_test)

train_control <- trainControl(method = "cv", number = 10)

logistic_fit <- train(Default ~ .,
                 data = macro_train,
                 trControl = train_control,
                 method = "glm",
                 family = "binomial")

logistic_fit_pred <- predict(logistic_fit, macro_test)

confusionMatrix(logistic_fit_pred, macro_test$Default)

postResample(pred = logistic_fit_pred, obs = macro_test$Default)

logistic_test <- mutate(macro_test,
                        obs = Default,
                        pred = predict(logistic_fit, newdata = macro_test),
                        predict(logistic_fit, newdata = macro_test, type = "prob"))

twoClassSummary(logistic_test, lev = levels(logistic_test$Default))
```



```{r}
tree = rpart(Default ~ ., data = macro_train)

min_cp <- tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]

tree_prune <- prune(tree, cp = min_cp)

tree_test_pred <- predict(tree, macro_test)

randomforest_grid_ranger <- expand.grid(mtry = seq(2, 4, length = 1), 
                                 splitrule = "extratrees",
                                 min.node.size = 2)

oob <- trainControl(method = "oob")

randomforest_tune <- train(Default ~ ., 
                           data = macro_train,
                           method = "ranger",
                           trControl = oob,
                           tuneGrid = randomforest_grid_ranger)

pred_randomforest_tune <- predict(randomforest_tune, macro_test)

confusionMatrix(pred_randomforest_tune, macro_test$Default)

postResample(pred_randomforest_tune, macro_test$Default)
```



```{r}
library(xgboost)
boosting <- train(Default ~ ., 
                  data = macro_train,
                  method = "xgbTree",
                  trControl = trainControl(method = "cv", number = 6)
)

pred_boosting <- predict(boosting, macro_test)

confusionMatrix(pred_boosting, macro_test$Default)

postResample(pred = pred_boosting, obs = macro_test$Default)

boosting_test <- mutate(macro_test,
                        obs = Default,
                        pred = predict(boosting, newdata = macro_test),
                        predict(boosting, newdata = macro_test, type = "prob"))

twoClassSummary(boosting_test, lev = levels(boosting_test$Default))
```
